### 2023-04-11 tesis_Luis/Chapter1/chapter1.tex
updated

In general, a particle detector is composed of a cylindrical or polygonal barrel section, with its axis aligned parallel to the incoming colliding beams. The cylindrical structure is sealed by two flat end caps, providing almost complete solid angle coverage down to the beam pipe, esta compuesto por varias capas dddicadas a detecta diferentes particulas.\\

@@

Colliding beam machines have the advantage that they can achieve much higher centre-of-mass energies, at various points the beams intersect and scattering takes place. If the incident particles have the same energy (equal mass) and therefore equal but opposite momentum, then the experiment is taking place in the centre-of-mass (CM) frame and the full energy delivered by the accelerator, can be used to produce high mass particles  \cite{undergraduate_accelerators_chapter}.
In particle physics the Lorentz invariant quantity $s$ denote the square of the total incoming energy and is defined as: 
\begin{equation}
s = \left ( \sum_{i = 1,2}^{}E_{i} \right )^{2}-\left ( \sum_{i = 1,2}^{}p_{i} \right )^{2}c^{2}
\end{equation}
Where $E$ and \textbf{$p$}  are the energy and the momenta of each incoming particle respectively, in the CM frame, where the momenta are equal and opposite, the second term vanishes and  $s = 4E^2$.

The second important parameter of the collider's performance is the luminosity, defined as the  quantity that measures the ability of a particle accelerator to produce the required number of interactions and is the proportionality factor  between the number of events of particular physics process per second $dR/dt$ and the quantum mechanical probability for the interaction called cross section \sigma_{p} \cite{ref_lib_vol3}.

%\section{Cross section}
%%resumir en un parrafo de 1.3 a 1.7
The technical meaning of "cross section" in particle physics is quite different from its common usage, while "cross section" typically refers to a slice of an object, in particle physics, it often refers to the probability that two particles will collide and react in a specific manner. When two beams cross in a particle accelerator, a variety of different processes can occur and  the cross section of a given process depends on the type and energy of the colliding particles.
At the Large Hadron Collider (LHC), certain particles such as  $W^{\pm}$ and $Z^{0}$ bosons have large cross sections , making their observation more frequent, while the production of a Higgs boson has a much lower cross section, making it more difficult to produce \cite{thomson_2013}.\\

@@

Where the value of $\mathcal{L}{\text{inst}}$ is evaluated over time $T$ (excluding possible dead time) because its intensity decreases as the number of protons reduces during the collisions of the fill. So $\mathcal{L}{int}$ is directly related to the number of observed events as:

@@

Luminosity is one of the fundamental parameters used to measure the performance of an accelerator, understanding luminosity with the best possible precision is crucial for physics analyses such as searches for new particles, rare processes, or measurements of known particle properties \cite{lumi_motiv}.
The experimental techniques to determine signal rates are mature enough, where the understanding of acceptances, detector biases, reconstruction efficiencies, or background subtraction is at the sub-percent level, so that the final precision of the physics measurement is dominated by the luminosity uncertainty \cite{lumi_paper_def_and_concept}. Ultimately, the greater the precision of the luminosity measurement, the more physicists can understand their observations and delve into hidden corners beyond our current knowledge.
%Knowing the integrated luminosity enables physicists to compare their observations with theoretical predictions and simulations. For instance, they can search for dark matter particles that escape collisions undetected by examining the energies and momenta of all particles generated in a collision. If there is an imbalance, it could be due to an undetected, potentially dark matter, particle that carries energy away. This is a powerful method of searching for new phenomena, but many effects, such as neutrinos produced in collisions, need to be considered. Neutrinos also escape undetected and leave an energy imbalance, making them indistinguishable from new phenomena. Therefore, to determine if something unexpected has been produced, physicists must look at the numbers \cite{luminosity_importance} .

%For example, if 11,000 events show an energy imbalance, and the simulations predict 10,000 events containing neutrinos, this could be significant. But if physicists only know the luminosity with a precision of 10\%, they could have easily had 11,000 neutrino events, and there were just 10\% more collisions than assumed. Clearly, a precise determination of luminosity is critical   \cite{luminosity_importance}.

%There are also types of analyses that depend much less on absolute knowledge of the number of collisions. For example, in measurements of ratios of different particle decays, such as the recent LHCb measurement, uncertainties in luminosity get canceled out in the ratio calculations. Other searches for new particles look for peaks in mass distribution and rely more on the shape of the observed distribution and less on the absolute number of events. But these also need to know the luminosity for any interpretation of the results \cite{luminosity_importance}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Expected  Luminosity precision for LHC Run 3}

Since its inception in 2010 the LHC has undergone three runs and two lengthy shutdowns for maintenance and upgrades. 
Run1 began in 2010 with proton-proton collisions at a center-of-mass energy of $\sqrt{s}=\text{7 TeV}$ and concluded in August 2012 at  $\sqrt{s}=\text{8 TeV}$\cite{LHC_status_2013}. Run 2, which took place at $\sqrt{s}=\text{13 TeV}$, was divided into two stages; the first  occurred from 2015 to 2016, and the second part  was carried out   from 2017 to 2018 \cite{pas_17,pas_18}, efforts are currently underway to further improve the precision of this run. Finally  Run3, which began in 2022 at $\sqrt{s}=\text{13.6 TeV}$ and is currently ongoing. This thesis project aims to contribute to the improvement of the precision of the luminosity for Run 3, where it is expected to obtain an approximate luminosity of $390 fb^{-1}$.

@@

\begin{table}[h!]
    \begin{center}
    \caption{Integrated luminosity values recorder by each Run}
	\label{lumi_values}
\begin{tabular}{cccc}
\textbf{Run} & \textbf{Energy $\sqrt{s}$} & \begin{tabular}[c]{@{}c@{}}\textbf{Integrated}\\\textbf{Luminosity}\end{tabular} & \textbf{Precision}    \\ 
\toprule
1(2010-2012) & 8 TeV           & $30 fb^{-1}$                                                                               & 2.5\%                 \\
2(2015-2016) & 13 TeV          & $45.9 fb^{-1}$                                                                             & 1.3\%                 \\
2(2017-2018) & 13 TeV          & $118 fb^{-1}$                                                                              & 2.5\%                 \\
3(2022-2024) & 13.6 TeV        & \begin{tabular}[c]{@{}c@{}}$45 fb^{-1}$(current)\\$390 fb^{-1}$(expected)\end{tabular}                  & \begin{tabular}[c]{@{}c@{}}currently\\estimating  \end{tabular}\\
\end{tabular}
    \end{center}
\end{table}

\begin{table}[h!]
\begin{center}
    \caption{Integrated luminosity values recorder by each Run}
	\label{lumi_values2}
\begin{tabular}{ccccc}
\multicolumn{1}{l}{\textbf{Run}} & \textbf{Year} & \textbf{Energy $\sqrt{s}$} & \begin{tabular}[c]{@{}c@{}}\textbf{Integrated}\\\textbf{Luminosity}\end{tabular} & \textbf{Precision}    \\ 
\toprule
\multirow{3}{*}{1}               & 2010          & 7 TeV           & $45 pb^{-1}$                                                                                 & 2.5\%                 \\
                                 & 2011          & 7 TeV           & $6.1 fb^{-1}$                                                                                & 2.4\%                 \\
                                 & 2012          & 8 TeV           & $23.3 fb^{-1}$                                                                               & 2.5\%                 \\ 
\hline
\multirow{2}{*}{2}               & 2015          & 13 TeV          & $4.3 fb^{-1}$                                                                              & 1.6\%                 \\
                                 & 2016          & 13 TeV          & $2.6 fb^{-1}$                                                                              & 1.2\%                 \\ 
\hline
\multirow{2}{*}{2}               & 2017          & 13 TeV          & $49.8 fb^{-1}$                                                                             & 2.3\%                 \\
                                 & 2018          & 13 TeV          & $67.9 fb^{-1}$                                                                              & 2.5\%                 \\ 
\hline
\multirow{2}{*}{3}               & 2022          & 13.6 TeV        & $45 fb^{-1}$                                                                                & currently estimating  \\
                                 & 2023          & 13.6 TeV        & currently recording                                                              & currently estimating 
\end{tabular}
 \end{center}
\end{table}

\begin{center}
  \begin{figure}[h!]
    \centering
    \includegraphics[scale=.28]{Chapter1/int_lumi_per_year.png}
    \caption[CMS Luminosity per year]{Delivered luminosity versus time for Run-1 (2010-2012)\cite{pas_17} and Run-2 (2015-2018)\cite{pas_18} and Run-3 that is still ongoing . Cumulative luminosity versus day delivered to CMS during stable beams for pp collisions at nominal center-of-mass energy.  \cite{wikicern}.}
    \label{lumi_per_year_int}
  \end{figure}
\end{center}
%% tabla con ano , luminosidad y presicion

>>>>

@@

Colliding beam machines have the advantage that they can achieve much higher centre-of-mass energies, at various points the beams intersect and scattering takes place. If the incident particles have the same energy (equal mass) and therefore equal but opposite momentum, then the experiment is taking place in the centre-of-mass (CM) frame and the full energy delivered by the accelerator, can be used to produce high mass particles  \cite{undergraduate_accelerators_chapter}.
In particle physics the Lorentz invariant quantity $s$ denote the square of the total incoming energy and is defined as: 
\begin{equation}
s = \left ( \sum_{i = 1,2}^{}E_{i} \right )^{2}-\left ( \sum_{i = 1,2}^{}p_{i} \right )^{2}c^{2}
\end{equation}
Where $E$ and \textbf{$p$}  are the energy and the momenta of each incoming particle respectively, in the CM frame, where the momenta are equal and opposite, the second term vanishes and  $s = 4E^2$.\\

The second important parameter of the collider's performance is the luminosity, defined as the  quantity that measures the ability of a particle accelerator to produce the required number of interactions and is the proportionality factor  between the number of events of particular physics process per second $dR/dt$ and the quantum mechanical probability for the interaction called cross section \sigma_{p} \cite{ref_lib_vol3}.\\

The technical meaning of "cross section" in particle physics is quite different from its common usage, while "cross section" typically refers to a slice of an object, in particle physics, it often refers to the probability that two particles will collide and react in a specific manner. When two beams cross in a particle accelerator, a variety of different processes can occur and  the cross section of a given process depends on the type and energy of the colliding particles.
At the Large Hadron Collider (LHC), certain particles such as  $W^{\pm}$ and $Z^{0}$ bosons have large cross sections , making their observation more frequent, while the production of a Higgs boson has a much lower cross section, making it more difficult to produce \cite{thomson_2013}.\\

@@

Where the value of $\mathcal{L}_{inst}$ is evaluated over time $T$ (excluding possible dead time) because its intensity decreases as the number of protons reduces during the collisions of the fill. So $\mathcal{L}_{int}$ is directly related to the number of observed events as:

@@

Understanding luminosity with the best possible precision is crucial for physics analyses such as searches for new particles, rare processes, or measurements of known particle properties \cite{lumi_motiv}.
The experimental techniques to determine signal rates are mature enough, where the understanding of acceptances, detector biases, reconstruction efficiencies, or background subtraction is at the sub-percent level, so that the final precision of the physics measurement is dominated by the luminosity uncertainty \cite{lumi_paper_def_and_concept}. Ultimately, the greater the precision of the luminosity measurement, the more physicists can understand their observations and delve into hidden corners beyond our current knowledge.

@@

\section{Expected  Luminosity precision for LHC Run 3}

Since its inception in 2010 the LHC has undergone three runs and two lengthy shutdowns for maintenance and upgrades. 
Run1 began in 2010 with proton-proton collisions at a center-of-mass energy of $\sqrt{s}=\text{7 TeV}$ and concluded in August 2012 at  $\sqrt{s}=\text{8 TeV}$\cite{LHC_status_2013}. Run 2, which took place at $\sqrt{s}=\text{13 TeV}$, was divided into two stages; the first  occurred from 2015 to 2016, and the second part  was carried out   from 2017 to 2018 \cite{pas_17,pas_18}, efforts are currently underway to further improve the precision of this run. Finally  Run3, which began in 2022 at $\sqrt{s}=\text{13.6 TeV}$ and is currently ongoing. This thesis project aims to contribute to the improvement of the precision of the luminosity for Run 3, where it is expected to obtain an approximate luminosity of $390 fb^{-1}$. \\

@@

\begin{table}[h!]
    \begin{center}
    \caption{Integrated luminosity values recorder by each Run}
	\label{lumi_values}
\begin{tabular}{cccc}
\textbf{Run} & \textbf{Energy $\sqrt{s}$} & \begin{tabular}[c]{@{}c@{}}\textbf{Integrated}\\\textbf{Luminosity}\end{tabular} & \textbf{Precision}    \\ 
\toprule
1(2010-2012) & 8 TeV           & $30 fb^{-1}$                                                                               & 2.5\%                 \\
2(2015-2016) & 13 TeV          & $45.9 fb^{-1}$                                                                             & 1.3\%                 \\
2(2017-2018) & 13 TeV          & $118 fb^{-1}$                                                                              & 2.5\%                 \\
3(2022-2024) & 13.6 TeV        & \begin{tabular}[c]{@{}c@{}}$45 fb^{-1}$(current)\\$390 fb^{-1}$(expected)\end{tabular}                  & \begin{tabular}[c]{@{}c@{}}currently\\estimating  \end{tabular}\\
\end{tabular}
    \end{center}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{comment}
\begin{table}[h!]
\begin{center}
    \caption{Integrated luminosity values recorder by each Run}
	\label{lumi_values2}
\begin{tabular}{ccccc}
\multicolumn{1}{l}{\textbf{Run}} & \textbf{Year} & \textbf{Energy $\sqrt{s}$} & \begin{tabular}[c]{@{}c@{}}\textbf{Integrated}\\\textbf{Luminosity}\end{tabular} & \textbf{Precision}    \\ 
\toprule
\multirow{3}{*}{1}               & 2010          & 7 TeV           & $45 pb^{-1}$                                                                                 & 2.5\%                 \\
                                 & 2011          & 7 TeV           & $6.1 fb^{-1}$                                                                                & 2.4\%                 \\
                                 & 2012          & 8 TeV           & $23.3 fb^{-1}$                                                                               & 2.5\%                 \\ 
\hline
\multirow{2}{*}{2}               & 2015          & 13 TeV          & $4.3 fb^{-1}$                                                                              & 1.6\%                 \\
                                 & 2016          & 13 TeV          & $2.6 fb^{-1}$                                                                              & 1.2\%                 \\ 
\hline
\multirow{2}{*}{2}               & 2017          & 13 TeV          & $49.8 fb^{-1}$                                                                             & 2.3\%                 \\
                                 & 2018          & 13 TeV          & $67.9 fb^{-1}$                                                                              & 2.5\%                 \\ 
\hline
\multirow{2}{*}{3}               & 2022          & 13.6 TeV        & $45 fb^{-1}$                                                                                & currently estimating  \\
                                 & 2023          & 13.6 TeV        & currently recording                                                              & currently estimating 
\end{tabular}
 \end{center}
\end{table}
\end{comment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
  \begin{center}
  \begin{figure}[ht]
    \centering
    \includegraphics[scale=.28]{Chapter1/int_lumi_per_year.png}
    \caption[CMS Luminosity per year]{Delivered luminosity versus time for Run-1 (2010-2012)\cite{pas_17} and Run-2 (2015-2018)\cite{pas_18} and Run-3 that is still ongoing . Cumulative luminosity versus day delivered to CMS during stable beams for pp collisions at nominal center-of-mass energy.  \cite{wikicern}.}
    \label{lumi_per_year_int}
  \end{figure}
    \end{center}
%% tabla con ano , luminosidad y presicion

### 2023-04-11 tesis_Luis/Chapter3/chapter3.tex
updated

\chapter{Luminosity  Measurement and Calibration}

@@

\subsection*{Backgrounds}

>>>>

\chapter{Luminosity  Measurement and Calibration}
\label{ch3}

Accurately measuring the luminosity delivered to the CMS experiment by the LHC is essential for various reasons. Online, the luminosity measurement provides feedback on the LHC and CMS performance and operations, including measuring trigger rates. In offline analysis, the luminosity measurement is a critical component for measuring the cross-section of observed processes or setting upper limits in searches for processes beyond the standard model.

As we see, to measure luminosity, a total of seven luminometers are used at CMS, and each of them reads out a specific quantity observed in the detector, such as hits, tracks, or clusters. The rate R measured by the luminometer is proportional to the instantaneous luminosity, $\mathcal{L}_{inst}$, with the constant of proportionality given by the visible cross-section $\sigma_{vis}$ \cite{pas_18}.

\begin{equation}
R(t)=\mathcal{L}_{inst}\sigma_{vis}
\label{lumi_exp_gen}
\end{equation}

The determination of $\sigma_{vis}$ is performed through van der Meer (vdM) scans using a specialized LHC machine setup. This chapter provides a description of this procedure, specifically applied with the PCC method.

\section{Pixel Cluster Counting method}

The PCC method utilizes the mean number of pixel clusters in the CMS silicon pixel detector to determine an offline luminosity measurement, this method exploits the very large number of pixels in the inner part of the CMS tracking system, it means that the probability of a given pixel being hit by two different charged particles from the same bunch crossing is exceedingly small, this large granularity and low occupancy yield measurements with excellent linear response with the pile-up (number of interactions per bunch crossing)$\mu$ \cite{PCC_PAS_12_001},  on average, the detector occupancy is less than 0.1\%\cite{lumi_precise_2015_2016}.\\

Figure  \ref{pileup} shows a simulation with pp collision zero-bias events with a full CMS detector response made in 2016, it contain only in time pileup and detector noise, where the mean number of pixel clusters is of the order of  100 per event. For simplicity, the number of pileup interactions present in each event is randomly generated from a representative PCC distribution at $\mu$ = 45 and the average PCC as a function of $\mu$. The latter distribution is fitted with a first-order polynomial, assuming no correlations among different values of $\mu$. Good agreement is seen based on the estimated goodness-of-fit $\chi^{2}$ per degree of freedom (dof) value of about 0.5 , indicating linearity under simulated conditions \cite{lumi_precise_2015_2016}.

\begin{center}
  \begin{figure}[h!]
    \centering
    \includegraphics[scale=.3]{Chapter3/pileup_lineality.png} 
    \caption[Linearity with pile-up]{ The left plot shows the number of pixel clusters and their statistical spread from simulation of pileup following a Poisson distribution with a mean of 45. The right plot shows the mean number of pixel clusters from simulation as a function of mean pileup. The red curve is a first-order polynomial fit with slope and $\chi^{2}/ndof$ values shown in the legend. The lower panel of the right plot shows the difference between the simulation and the linear fit in black points. The green band is the final linearity uncertainty for the 2016 data set \cite{lumi_precise_2015_2016} }
    \label{pileup}
  \end{figure}
\end{center}

To obtain the mean number of pixel clusters per event, several zero-bias events are averaged. This value is given by :

\begin{equation}
\left < N_{\text{cluster}} \right > \equiv \left < N_{\text{cluster}/\text{interaction}} \right > \mu
\end{equation}

Under minimum bias conditions, $\mu$ can be expressed as:

\begin{equation}
\mu = \frac{\sigma_{\text{minBias}}}{f} \cdot \mathcal{L}_{inst}
\end{equation}

where $f$ is the LHC revolution frequency and $\mathcal{L}_{inst}$ is the single bunch instantaneous luminosity (SBIL). The minimum bias cross section $\sigma_{\text{minBias}}$ here is related to the PCC visible cross section by the mean number of clusters per interaction:

\begin{equation}
\sigma_{vis}= \left < N_{\text{cluster}/\text{interaction}} \right >\cdot \sigma_{\text{minBias}}
\end{equation}

Combining everything, the PCC luminosity measurement is obtained as:

\begin{equation}
\mathcal{L}_{inst}=\frac{\left < N_{\text{cluster}} \right >}{\sigma_{vis}} \cdot f
\end{equation}

This method is capable of providing a precise luminosity measurement over longer time periods, but it cannot do so for a single luminosity section (23 seconds) as is possible with online luminometers. The reason for this limitation is the limited CMS trigger bandwidth available for collecting data \cite{pas_18}. A detailed discussion on this topic can be found in the next chapter.\\

In the PCC measurement, the innermost layer of the pixel detector is excluded from the analysis due to significant dynamic inefficiency effects \cite{pas_18}. At higher $\mathcal{L}_{inst}$, the hit efficiency in this layer decreases because the readout chip cannot process all of the hits.\\

\section{2022 CMS Data collected}

On July 12th, 2022, the data collection for Run3 of the Large Hadron Collider (LHC) began, which resulted in a total luminosity of $38.48 fb^{-1}$. This collection period encompassed the duration from the start of stable beams until the beam was dumped, or when stable beams ceased for other reasons, such as beam studies. The luminosity delivered by the LHC to CMS (blue) and the luminosity recorded by CMS (orange) during stable beams and proton-proton collisions at a center-of-mass energy of $\sqrt{s}$ 13.6 TeV in 2022 are depicted in Figure \ref{Lumi_2022}. Note that  the special runs that occurred at injection energy (450 GeV) are not represented in these graphs.

\begin{center}
  \begin{figure}[h!]
    \centering
    \includegraphics[width=.45\textwidth]{Chapter3/lumi_per_day_cumulative_pp_2022.png}
    \includegraphics[width=.45\textwidth]{Chapter3/lumi_per_day_pp_2022.png}
    \caption[Integrated Luminosity 2022]{(left) Cumulative day-by-day integrated luminosity, 2022 Cumulative measured luminosity versus day delivered. Note that this plot may not include data taken within the most recent 24 hours. (rigth) Day-by-day integrated luminosity, 2022 As the first plot, but not cumulative.} 
    \label{Lumi_2022}
  \end{figure}
\end{center}

\section{Module selection}

To ensure accurate luminosity measurement, a module selection is created to remove any modules exhibiting long-term performance variations, indicating a non-physical shift in their cluster counts. With a total of 1856 modules in the pixel system, all of them can be considered in the luminosity measurement; however, non-zero occupancy and non-linear effects can pose challenges for accurate measurement. Thus, to identify "good" and "stable" modules, a subset is selected by eliminating  underperforming or "bad" modules and comparing their relative contributions to the cross section. Those with relatively consistent contributions across standard physics runs are kept, while those with significant changes in their relative cross section compared to the averaged relative contribution are rejected. The Module selection is achieved  with the following procedure:

\begin{itemize}
\item Poor statistics lumisections are removed by applying selection on total PCC.
\item Barrel layer-1 modules are removed, as these modules are significantly affected by dynamic inefficiency.
\item A loose selection of 7\% based on RMS/mean frome the module weight distribution (changes in the module PCC ratio relative to the total PCC) is applied. Modules with significantly large RMS/mean are removed with this selection as shown in \ref{goodmodule}.
\item The module stability is re-evaluated based on RMS/mean values using an iterative method where the second iteration is made with a loose selection of 4\%.
\end{itemize}

\begin{center}
  \begin{figure}[h!]
    \centering
    \includegraphics[scale=.17]{Chapter4/good_module.png}\includegraphics[scale=.15]{Chapter4/RMSmean.png}
    \caption[RMS/mean Module  Stability]{ left: stability profile for good and bad modules where weight is plotted as a function of lumi section.  Right: RMS/mean values of module weight for all pixel modules  where each pixel module is represented by a module number index (same happen for the 4\% iteration).} 
    \label{goodmodule}
  \end{figure}
\end{center}

The changing conditions of the pixel detector over time, such as detector noise, aging effects, and radiation damage, can cause variations in the number of well-performing modules during different periods. Recognizing this, these variations are evaluated over a 23-second interval, which corresponds to the granularity of the luminosity database.\\

To ensure the accuracy of the measurements, pixel module veto lists were generated for five different time periods, namely C, D, E, F, and G, using the procedure described above for each period. To further enhance the stability of the PCC measurement, a common module vetolist was created by combining the lists from all five periods were the number of bad modules per period were added and modules with equal ID were removed , as shown in table \ref{common veto module}. The PCC data with zero bias was then reprocessed using this common module vetolist and ultimately used for the VdM analysis.

\begin{table}[ht]
\centering
\caption{After each iteration for each period, the periods C and D are combined, followed by the periods C, D, and E, then the periods C, D, E, and F, and finally, the common veto list is created using the periods C, D, E, F, and G.}
\begin{tabular}{ccc}
\textbf{Period} & \textbf{\# bad modules} & \textbf{\# good modules}  \\ 
\toprule
2022C+D         & 886                     & 970                       \\
2022C+D+E       & 1106                    & 750                       \\
2022C+D+E+F     & 1307                    & 549                       \\
2022C+D+E+F+G   & 1411                    & 445                   
\label{common  veto module}   
\end{tabular}
\end{table}

@@

As discussed in Chapter 1, the instantaneous luminosity for a single colliding bunch is described by Eq. (\ref{luminosity_2}). In practice, while the measurement of the beam currents $N_{1,2}$ is well determined, the individual proton density functions cannot be directly measured. To address this, the VdM method involves a specific machine setup that allows for the determination of the two beam overlap integrals. This is achieved by varying the separation between the beams and measuring the resulting rates, which can be used to obtain density profiles that are close to normal distributions.

\begin{equation}
\int \rho_{x1}(x) \rho_{x2}(x) dx = \frac{R_{x}(0)}{\int R_{x}(\Delta) d\Delta}
\end{equation}

where $R_{x}(\Delta)$ is the rate measured when the two beams are separated in $x$ by a distance $\Delta$; a asimilar equation can be written in $y$. Then the beam overlap width $\Sigma_{x}$ (and similarly $\Sigma_{y}$) is defined as \cite{pas_18}:

\begin{equation}
\Sigma_{x}= \frac{1}{\sqrt{2\pi}} \frac{\int R_{x}(\Delta)d\Delta}{R_{x}(0)}
\label{CapSigma}
\end{equation}

This process leads to the final expression for luminosity for a single colliding bunch:

\begin{equation}
\mathcal{L}_{inst}=\frac{N_{1} N_{2}f}{2 \pi \Sigma_{x}\Sigma_{y}}
\end{equation}

where $N_{1,2}$ are the particles per bunch (bunch current) and  $f= 11246$ Hz is the bunch orbit frequency around the LHC ring determined by the energy of the protons.\\

Therefore, the equation \ref{lumi_exp_gen} used to measure the visible cross sections $\sigma_{vis}$ takes the following form:

\begin{equation}
  \sigma_{vis}=\frac{2\pi \Sigma_{x} \Sigma_{y} R(0, 0)}{N_{1}N_{2} f}
  \label{sigmavis_eq}
\end{equation}

Experimentally, the quantities $\Sigma_{x}$ and $\Sigma_{y}$, as defined in Eq. (\ref{CapSigma}), are determined by conducting two separate scans in the $x$ and $y$ directions, respectively. 
These scans are performed by varying the separation between the beams in each direction  at a fixed number of separation steps.The separation steps are determined through curve fitting of the scan data based on the luminometer rate measurements obtained during the vdM scans.\\

The measured rate, denoted by $R(0,0)$, is determined as the highest value among the rates obtained in the $x$-scan $R_x(0)$, and the rates obtained in the $y$-scan $R_y(0)$. Although the beam widths are identical for all luminometers, the peaks of the scan curves are specific to each luminometer.\\

Figure \ref{vdm_sketch} depicts a schematic of the beam positions during vdM scans in the $x$ and $y$ planes, along with the detector rate as a function of beam separation  \cite{pas_18}.

\begin{center}
  \begin{figure}[h!]
    \centering
    \includegraphics[scale=.37]{Chapter3/vdm_sketch.png}
    \caption[Sketch of a vdM scan in x and y planes and example of fitting resulting rates]{ The sketch of a vdM scan in $x$ and $y$ planes. The indent sketch is an example of the fit of the resulting rates \cite{vdM_sketch}.}
    \label{vdm_sketch}
  \end{figure}
\end{center}
